---
editor_options: 
  chunk_output_type: console
---

# Extent of elevational migration

In this script, we calculate the median migration extent, breeding and wintering elevation for all combinations of elevayional limit and sampling effort-seperately for the eastern and western Himalayas

## Load necessary libraries
```{r}
library(ape)
library(geiger)
library(phytools)
library(coxme)
library(evobiR)
library(nlme)
library(tidyverse)
library(phangorn)
```

## Eastern Himalayas
```{r}
# load previously generated .Rdata files for analyses

for (minSample in c(30, 60)) {
  for (ds in c("05.q1", "05.q2", "05.q3", "95.q1", "95.q2", "95.q3", "50.q1", "50.q2", "50.q3")) {
    load(paste("eBird_resampled_east_", ds, ".Rdata", sep = ""))
    B.elev <- sapply(1:1000, function(x) rs[[x]][, 1])
    B.n <-    sapply(1:1000, function(x) rs[[x]][, 2])
    W.elev <- sapply(1:1000, function(x) rs[[x]][, 3])
    W.n <-    sapply(1:1000, function(x) rs[[x]][, 4])
    
    diff <- B.elev - W.elev
    dimnames(diff)<-list(uniSpe[,1],1:1000)
    
    # For each species exclude trials where it has been detected less than 30/60 times in winter or summer
    if (minSample == 30) diff[W.n < 30 | B.n < 30] <- NA
    if (minSample == 60) diff[W.n < 60 | B.n < 60] <- NA 
    diff.sel <- diff[apply(diff, 1, FUN = function(x) sum(is.na(x))) < 1000, ]
    
    # Calculate medians of the percentiles over 1000 sets of resampled sampling events. Along with confidence intervals    
    med.CI <- apply(diff.sel, 1, FUN = function(x) quantile((x), c(0.025, .5, .975), na.rm = TRUE))
    med.CI <- t(med.CI)
    med.CI  <- as.data.frame(med.CI)
    med.CI$Group <- with(med.CI, ifelse(`2.5%` > 0 & `97.5%` > 0, 2, ifelse(`2.5%` < 0 & `97.5%` < 0, 1, 0)))
    med.CI$LCI_diff <- as.numeric(med.CI$`2.5%`)
    med.CI$Median_diff <- as.numeric(med.CI$`50%`)
    med.CI$UCI_diff <- as.numeric(med.CI$`97.5%`)
    med.CI$Species <- rownames(med.CI)
    med.CI<-med.CI[,-c(1,2,3)]
    
    ## adding breeding elevations to the same table
    
    dimnames(B.elev)<-list(uniSpe[,1],1:1000)
    if (minSample == 30) B.elev[W.n < 30 | B.n < 30] <- NA
    if (minSample == 60) B.elev[W.n < 60 | B.n < 60] <- NA 
    B.sel <- B.elev[apply(B.elev, 1, FUN = function(x) sum(is.na(x))) < 1000, ]
    
    med.S.CI <- apply(B.sel, 1, FUN = function(x) quantile((x), c(0.025, .5, .975), na.rm = TRUE))
    med.S.CI <- t(med.S.CI)
    med.S.CI  <- as.data.frame(med.S.CI)
    med.S.CI$LCI_S <- as.numeric(med.S.CI$`2.5%`)
    med.S.CI$Median_S <- as.numeric(med.S.CI$`50%`)
    med.S.CI$UCI_S <- as.numeric(med.S.CI$`97.5%`)
    med.S.CI$Species <- rownames(med.S.CI)
    med.S.CI<-med.S.CI[,-c(1,2,3)]
    
    ## adding winter elevations to the same table
    
    dimnames(W.elev)<-list(uniSpe[,1],1:1000)
    if (minSample == 30) W.elev[W.n < 30 | B.n < 30] <- NA
    if (minSample == 60) W.elev[W.n < 60 | B.n < 60] <- NA 
    W.sel <- W.elev[apply(W.elev, 1, FUN = function(x) sum(is.na(x))) < 1000, ]
    
    med.W.CI <- apply(W.sel, 1, FUN = function(x) quantile((x), c(0.025, .5, .975), na.rm = TRUE))
    med.W.CI <- t(med.W.CI)
    med.W.CI  <- as.data.frame(med.W.CI)
    med.W.CI$LCI_W <- as.numeric(med.W.CI$`2.5%`)
    med.W.CI$Median_W <- as.numeric(med.W.CI$`50%`)
    med.W.CI$UCI_W <- as.numeric(med.W.CI$`97.5%`)
    med.W.CI$Species <- rownames(med.W.CI)
    med.W.CI<-med.W.CI[,-c(1,2,3)]
    
    ts<-left_join(med.CI,med.S.CI, by = "Species") %>% left_join(., med.W.CI, by = "Species", )
    
    ts <- ts[c("Species", "Group", "Median_S", "LCI_S", "UCI_S","Median_W","LCI_W","UCI_W","Median_diff","LCI_diff", "UCI_diff")]
    
    write.csv(ts, file = paste("birdlist_east_",ds,".sam",minSample, ".csv", sep = ""), row.names = F)
  }
}
```

## Western Himalayas
```{r}
for (minSample in c(30, 60)) {
  for (ds in c("05.q1", "05.q2", "05.q3", "95.q1", "95.q2", "95.q3", "50.q1", "50.q2", "50.q3")) {
    load(paste("eBird_resampled_west_", ds, ".Rdata", sep = ""))
    B.elev <- sapply(1:1000, function(x) rs[[x]][, 1])
    B.n <-    sapply(1:1000, function(x) rs[[x]][, 2])
    W.elev <- sapply(1:1000, function(x) rs[[x]][, 3])
    W.n <-    sapply(1:1000, function(x) rs[[x]][, 4])
    
    diff <- B.elev - W.elev
    dimnames(diff)<-list(uniSpe[,1],1:1000)
    
    # For each species exclude trials where it has been detected less than 30/60 times in winter or summer
    if (minSample == 30) diff[W.n < 30 | B.n < 30] <- NA
    if (minSample == 60) diff[W.n < 60 | B.n < 60] <- NA 
    diff.sel <- diff[apply(diff, 1, FUN = function(x) sum(is.na(x))) < 1000, ]
    
    # Calculate medians of the percentiles over 1000 sets of resampled sampling events. Along with confidence intervals    
    med.CI <- apply(diff.sel, 1, FUN = function(x) quantile((x), c(0.025, .5, .975), na.rm = TRUE))
    med.CI <- t(med.CI)
    med.CI  <- as.data.frame(med.CI)
    med.CI$Group <- with(med.CI, ifelse(`2.5%` > 0 & `97.5%` > 0, 2, ifelse(`2.5%` < 0 & `97.5%` < 0, 1, 0)))
    med.CI$LCI_diff <- as.numeric(med.CI$`2.5%`)
    med.CI$Median_diff <- as.numeric(med.CI$`50%`)
    med.CI$UCI_diff <- as.numeric(med.CI$`97.5%`)
    med.CI$Species <- rownames(med.CI)
    med.CI<-med.CI[,-c(1,2,3)]
    
    ## adding breeding elevations to the same table
    
    dimnames(B.elev)<-list(uniSpe[,1],1:1000)
    if (minSample == 30) B.elev[W.n < 30 | B.n < 30] <- NA
    if (minSample == 60) B.elev[W.n < 60 | B.n < 60] <- NA 
    B.sel <- B.elev[apply(B.elev, 1, FUN = function(x) sum(is.na(x))) < 1000, ]
    
    med.S.CI <- apply(B.sel, 1, FUN = function(x) quantile((x), c(0.025, .5, .975), na.rm = TRUE))
    med.S.CI <- t(med.S.CI)
    med.S.CI  <- as.data.frame(med.S.CI)
    med.S.CI$LCI_S <- as.numeric(med.S.CI$`2.5%`)
    med.S.CI$Median_S <- as.numeric(med.S.CI$`50%`)
    med.S.CI$UCI_S <- as.numeric(med.S.CI$`97.5%`)
    med.S.CI$Species <- rownames(med.S.CI)
    med.S.CI<-med.S.CI[,-c(1,2,3)]
    
    ## adding winter elevations to the same table
    
    dimnames(W.elev)<-list(uniSpe[,1],1:1000)
    if (minSample == 30) W.elev[W.n < 30 | B.n < 30] <- NA
    if (minSample == 60) W.elev[W.n < 60 | B.n < 60] <- NA 
    W.sel <- W.elev[apply(W.elev, 1, FUN = function(x) sum(is.na(x))) < 1000, ]
    
    med.W.CI <- apply(W.sel, 1, FUN = function(x) quantile((x), c(0.025, .5, .975), na.rm = TRUE))
    med.W.CI <- t(med.W.CI)
    med.W.CI  <- as.data.frame(med.W.CI)
    med.W.CI$LCI_W <- as.numeric(med.W.CI$`2.5%`)
    med.W.CI$Median_W <- as.numeric(med.W.CI$`50%`)
    med.W.CI$UCI_W <- as.numeric(med.W.CI$`97.5%`)
    med.W.CI$Species <- rownames(med.W.CI)
    med.W.CI<-med.W.CI[,-c(1,2,3)]
    
    ts<-left_join(med.CI,med.S.CI, by = "Species") %>% left_join(., med.W.CI, by = "Species", )
    
    ts <- ts[c("Species", "Group", "Median_S", "LCI_S", "UCI_S","Median_W","LCI_W","UCI_W","Median_diff","LCI_diff", "UCI_diff")]
    
    write.csv(ts, file = paste("birdlist_west",ds,".sam",minSample, ".csv", sep = ""), row.names = F)
  }
}
```

![**Elevational shift in the eastern and western Himalayas**
Elevational shift in the eastern Himalayas (A) and western Himalayas (B). Figures show species estimates (at high sampling effort) of median summer elevations (orange circles) in descending order and their respective median winter elevations (teal triangles). Errors bars indicate 95% confidence intervals estimated from the resampling protocol.](figs/fig02.jpg)
