---
editor_options: 
  chunk_output_type: console
---

# Phylogenetic generalized least squares regressions

Here, we used phylogenetic generalized least squares regression to test if thermal regime, dispersal ability, and diet significantly drive Himalayan bird elevational shift.  

## Remove non-Himalayan species
```{r}
# first we remove species that are non-Himalayan
rem <-read.csv("data/SpeciesList_nonhimalayan.csv.csv")
```

## PGLS models for the Eastern Himalayas
```{r}
for (ds in c("05.q1.sam30", "05.q1.sam60", "05.q2.sam30","05.q2.sam60", "05.q3.sam30","05.q3.sam60", "95.q1.sam30","95.q1.sam60", "95.q2.sam30","95.q2.sam60" ,"95.q3.sam30","95.q3.sam60" ,"50.q1.sam30","50.q1.sam60" ,"50.q2.sam30","50.q2.sam60" ,"50.q3.sam30", "50.q3.sam60")) {
  x<-read.csv(paste("results/final_birdlist_east_", ds, ".csv", sep = ""))
  x<-x[!(x$Species %in% rem$Species),]
  write.csv(x,file = paste("results/birdlist_east",ds,".csv", sep = ""), row.names = F) 
}

# load necessary data for running statistical models 
east_100 <-read.csv("results/eastHim_100.csv") #temperature data
east_100_S <- east_100 %>% select(-contains("Jan")) %>% rename(elev_roundS = elev_round) #winter
east_100_W <- east_100 %>% select(-contains("June"))%>% rename(elev_roundW = elev_round) #summer
jetzname <-read.csv("data/for_PGLS_list3.csv") #taxonomy used by birdtree.org
sheard_trait <- read.csv("data/2020-sheard et al-species-trait-dat - speciesdata.csv") %>% #trait data set
  select(HWI,Diet,Tree.name)
tree <-read.nexus("data/birdtree.nex")
tree <- mcc(tree) ## maximum clade credibility tree


compile_eastmodels<-setNames(data.frame(matrix(ncol = 6, nrow = 0)), c("Variable", "Value", "Std.Error", "t-value", "p-value", "ds"))

for (ds in c("05.q1.sam30", "05.q1.sam60", "05.q2.sam30","05.q2.sam60", "05.q3.sam30","05.q3.sam60", "95.q1.sam30","95.q1.sam60", "95.q2.sam30","95.q2.sam60" ,"95.q3.sam30","95.q3.sam60" ,"50.q1.sam30","50.q1.sam60" ,"50.q2.sam30","50.q2.sam60" ,"50.q3.sam30", "50.q3.sam60")) {
  x<-read.csv(paste("results/birdlist_east", ds, ".csv", sep = ""))
  x$elev_roundS<-round(x$Median_S,-2)
  x$elev_roundW<-round(x$Median_W,-2)
  x<-left_join(x,east_100_S, by = "elev_roundS")
  x<-left_join(x,east_100_W, by = "elev_roundW")
  x$ThermalRegime<-x$maxTemp_June_mean-x$minTemp_Jan_mean
  x<-left_join(x,jetzname, by = "Species")
  x<-left_join(x, sheard_trait, by = "Tree.name")%>% mutate(Diet = fct_recode(Diet, "omnivore" = "nectar", "omnivore" = "scav", "seeds"="plants"))
  pruned.tree<-drop.tip(tree1,tree1$tip.label[-match(x$Tree.name, tree1$tip.label)])
  glsdata<-x[match(pruned.tree$tip.label, x$Tree.name),]
 pgls2<-gls(Median_diff~ThermalRegime+HWI+relevel(Diet, ref = "invertebrates"),correlation=corPagel(1,pruned.tree),data=glsdata)
  k<-summary(pgls2)
  outputk<-as.data.frame(k$tTable)
  outputk$dataset<-ds
  outputk$Variable<-rownames(outputk)
  compile_eastmodels<-rbind(compile_eastmodels,outputk)
  
  write.csv(glsdata, file = paste("results/east_ordered",ds,".csv", sep = ""), row.names = F)
  write.csv(compile_eastmodels, "results/compile_eastmodels.csv", row.names = F)
}

east_p<-read.csv("results/compile_eastmodels.csv")
east_p$adjP<-p.adjust(east_p$p.value, method = "fdr")

write.csv(east_p, "results/compile_eastmodels_padj.csv")
```

## PGLS models for the Western Himalayas
```{r}
for (ds in c("05.q1.sam30", "05.q1.sam60", "05.q2.sam30","05.q2.sam60", "05.q3.sam30","05.q3.sam60", "95.q1.sam30","95.q1.sam60", "95.q2.sam30","95.q2.sam60" ,"95.q3.sam30","95.q3.sam60" ,"50.q1.sam30","50.q1.sam60" ,"50.q2.sam30","50.q2.sam60" ,"50.q3.sam30", "50.q3.sam60")) {
  x<-read.csv(paste("results/final_birdlist_west", ds, ".csv", sep = ""))
  x<-x[!(x$Species %in% rem$Species),]
  write.csv(x,file = paste("results/birdlist_west",ds,".csv", sep = ""), row.names = F) ## final list after removing non himalayan species
}

west_100<-read.csv("results/westHim_100.csv")
west_100_S<-west_100 %>% select(-contains("Jan")) %>% rename(elev_roundS = elev_round)
west_100_W<-west_100 %>% select(-contains("June"))%>% rename(elev_roundW = elev_round)

compile_westmodels<-setNames(data.frame(matrix(ncol = 5, nrow = 0)), c("Value", "Std.Error", "t-value", "p-value", "ds"))

for (ds in c("05.q1.sam30", "05.q1.sam60", "05.q2.sam30","05.q2.sam60", "05.q3.sam30","05.q3.sam60", "95.q1.sam30","95.q1.sam60", "95.q2.sam30","95.q2.sam60" ,"95.q3.sam30","95.q3.sam60" ,"50.q1.sam30","50.q1.sam60" ,"50.q2.sam30","50.q2.sam60" ,"50.q3.sam30", "50.q3.sam60")) {
  x<-read.csv(paste("results/birdlist_west", ds, ".csv", sep = ""))
  x$elev_roundS<-round(x$Median_S,-2)
  x$elev_roundW<-round(x$Median_W,-2)
  x<-left_join(x,west_100_S, by = "elev_roundS")
  x<-left_join(x,west_100_W, by = "elev_roundW")
  x$ThermalRegime<-x$maxTemp_June_mean-x$minTemp_Jan_mean
  x<-left_join(x,jetzname, by = "Species")
  x<-left_join(x, sheard_trait, by = "Tree.name")%>% mutate(Diet = fct_recode(Diet, "omnivore" = "nectar", "omnivore" = "scav", "seeds"="plants"))
  pruned.tree<-drop.tip(tree1,tree1$tip.label[-match(x$Tree.name, tree1$tip.label)])
  glsdata<-x[match(pruned.tree$tip.label, x$Tree.name),]
  pgls2<-gls(Median_diff~ThermalRegime+HWI+relevel(Diet, ref = "invertebrates"),correlation=corPagel(1,pruned.tree),data=glsdata)
  k<-summary(pgls2)
  outputk<-as.data.frame(k$tTable)
  outputk$dataset<-ds
  outputk$Variable<-rownames(outputk)
  compile_westmodels<-rbind(compile_westmodels,outputk)
  
  write.csv(glsdata, file = paste("results/west_ordered",ds,".csv", sep = ""), row.names = F)
  write.csv(compile_westmodels, "results/compile_westmodels.csv", row.names = F)
}

west_p<-read.csv("results/compile_westmodels.csv")
west_p$adjP<-p.adjust(west_p$p.value, method = "fdr")

write.csv(east_p, "results/compile_westmodels_padj.csv")

```

## Phylogenetic paired t-test
```{r}
diff_ttest<-setNames(data.frame(matrix(ncol = 3, nrow = 0)), c("t", "df", "pvalue"))
niche_ttest<-setNames(data.frame(matrix(ncol = 3, nrow = 0)), c("t", "df", "pvalue"))

for (ds in c("05.q1.sam30", "05.q1.sam60", "05.q2.sam30","05.q2.sam60", "05.q3.sam30","05.q3.sam60", "95.q1.sam30","95.q1.sam60", "95.q2.sam30","95.q2.sam60" ,"95.q3.sam30","95.q3.sam60" ,"50.q1.sam30","50.q1.sam60" ,"50.q2.sam30","50.q2.sam60" ,"50.q3.sam30", "50.q3.sam60")) {
  west<-read.csv(paste("results/west_ordered",ds,".csv", sep = ""))
  east<-read.csv(paste("results/east_ordered",ds,".csv", sep = ""))
  commonW<-west[(west$Species %in% east$Species),]
  commonE<-east[(east$Species %in% west$Species),]
  common<-data.frame(Median_diffW = commonW$Median_diff,Median_diffE = commonE$Median_diff, ThermalRegimeW = commonW$ThermalRegime,ThermalRegimeE = commonE$ThermalRegime ,Tree.name = commonE$Tree.name)
  pruned.tree<-drop.tip(tree1,tree1$tip.label[-match(common$Tree.name, tree1$tip.label)])
  comdata<-common[match(pruned.tree$tip.label, common$Tree.name),]
  row.names(comdata)<-comdata$Tree.name
  a<-phyl.pairedttest(pruned.tree,comdata[,c(1,2)])
  b<-phyl.pairedttest(pruned.tree,comdata[,c(3,4)])
  a_diff_ttest<-data.frame(t = a$t, pvalue = a$P, df = a$df, dataset = ds)
  b_niche_ttest<-data.frame(t = b$t, pvalue = b$P, df = b$df, dataset = ds)
  diff_ttest<-rbind(diff_ttest,a_diff_ttest)
  niche_ttest<-rbind(niche_ttest,b_niche_ttest)
  
  write.csv(diff_ttest, "results/diff_ttest.csv", row.names = F)
  write.csv(niche_ttest, "results/niche_ttest.csv", row.names = F)
}
```
